---
title: "hw5"
author: "cyber-atlas"
date: "April 9, 2019"
output: html_document
---

```{r}
library('dplyr')
library('tidyverse')
library('ggplot2')
```

First, let's get the dataset
```{r}
#The datasets gets saved in as data
if (file.exists("alcohol.rds")) {
  data <- readRDS("alcohol.rds")
} else {
  url <- "https://xdaiisu.github.io/ds202materials/data/ames.csv"
  data <- readr::read_csv(url)
  saveRDS(data, file="alcohol.rds")
} 


#Take a look at the dataset
head(data)

```

2. 

Extract the geographic latitude and longtitude and store in StateLocation
```{r}
#Using a regex to speperate the latitude and longtitude into different columns 
data <- data %>% separate(col="Store Location", into=c("Address", "Lat", "Lon"), sep="([\\(\\,\\)])")  

head(data)
```
Check the variable types and change them

```{r}
data$Lon <- as.numeric(data$Lon)
data$Lat <- as.numeric(data$Lat)
data$"Category Name" <- as.factor(data$"Category Name")
data$`Vendor Name` <- as.factor(data$`Vendor Name`)
data$City <- as.factor(data$City)
#str(data)

head(data)

```
extract year month day and set them as numeric

```{r}
data = separate(data=data, col="Date", into = c("Month", "Day", "Year"), sep="\\/")

data$Month = as.numeric(data$Month)
data$Day = as.numeric(data$Day)
data$Year = as.numeric(data$Year)
str(data)

```

3 

Plot a scatter plot of lat and long locations
```{r}
#Plot all the  non NA points
ggplot(data=subset(data, !is.na(c(Lat, Lon))), aes(x=Lat, y=Lon))+
  geom_point()

```

provide a visual vreakdown of the liquor categy 
```{r}

#ggplot(data, aes(x=factor(data$`Category Name`), fill =  ))
ggplot(data, aes(x='Volume Sold (Liters)', fill = data$`Category Name` ))+
  geom_bar()+
  coord_polar("y")+
  theme(legend.position = 'none')

#unique(data$`Category Name`)
#After running unique we know the nuymber of categories

#ggplot(data = subset(data$`Category Name`, 1:20), aes(x='Volume Sold (Liters)', fill = data$`Category Name` ))+
  #coord_polar("y")+
 # theme(legend.position = 'none')
  

```

4

Find the daily sales liquor sales in Ames in 2018. Summarize number of sales, volume of liquor sold and amount of money spent. Plot volume sold by day (use a scatterplot of volume by day and facet by month). Describe any patterns.
Find the dates for ISU football home games in Fall 2018. Can you see a pattern? 
```{r}

dataq5 <- filter(data, Year==2018 & City =='Ames')

#dataq5

q5sale <- dataq5 %>% group_by(Month, Day) %>% summarise(SumSales = sum(`Sale (Dollars)`))

head(q5sale)

q5volume <- dataq5 %>% group_by(Month, Day) %>% summarize(Sumvol = sum(`Volume Sold (Liters)`))

head(q5volume)

ggplot(q5volume, aes(x=Day, y=Sumvol))+
  geom_point()+
  stat_smooth()+
  facet_wrap(~Month)

```
The amounts did not spike much on football games. Every weekend had a spike. Ames drank a lot most weekends
